# Data transformation

(Work in progress)

## Read Raw Data

```{r read raw daata}
library(lubridate)
library(tidyverse)
library(zoo)

nyc_event_listing_events <- read.csv("data/raw_data/nyc_parks_event_listing_events.csv", na.strings = "")
names(nyc_event_listing_events) <- c(
  "EvtId", "Title", "Dt", "ST",  # event_id, title, date, start_time
  "ET", "LcDs", "Ds", "Snp", # end_time, location_description, description, snippet
  "Ph", "Em", "CsFr", "CsDs", # phone, email, cost_free, cost_description
  "MtSe", "Url", "Ntc" # must_see, url, notice
)
nyc_event_listing_events <- nyc_event_listing_events %>%
  mutate(Dt = as.Date(Dt, format = "%m/%d/%Y"))

nyc_event_categories <- read.csv("data/raw_data/nyc_parks_event_listing_categories.csv", na.strings = "")
names(nyc_event_categories) <- c("EvtId", "Cat") # event_id, category

nyc_event_organizers <- read.csv("data/raw_data/nyc_parks_event_listing_organizers.csv", na.strings = "")
names(nyc_event_organizers) <- c("EvtId", "Org") # event_id, event_organizer

nyc_event_listing_locations <- read.csv("data/raw_data/nyc_parks_event_listing_locations.csv", na.strings = "")
names(nyc_event_listing_locations) <- c(
  "EvtId", #event_id
  "LcNm", #name
  "PkId", #park_id
  "Lat", #latitude
  "Long", #longitude
  "Addr", #address
  "Zip",
  "Bo", #borough
  "Accs" #accessible
)

nyc_parks_crime <- read.csv("data/derived_data/nyc_parks_crime_data_all.csv", na.strings = "")
names(nyc_parks_crime) <- c(
  "Park", #PARK
  "Boro", #BOROUGH
  "Size", #SIZE..ACRES
  "Cat", #CATEGORY
  "Mdr", #MURDER
  "Rape", #RAPE
  "Rob", #ROBBERY
  "Ass", #FELONY.ASSAULT
  "Burg", #BURGLARY
  "G_L", #GRAND.LARCENY
  "GL_M_V", #GRAND.LARCENY.OF.MOTOR.VEHICLE
  "Tot", #TOTAL
  "Qtr"#QUARTER
)

crime_quarters <- c(
  "2021_q3",
  "2021_q2",
  "2021_q1",
  "2020_q4",
  "2020_q3",
  "2020_q2",
  "2020_q1",
  "2019_q4",
  "2019_q3",
  "2019_q2",
  "2019_q1",
  "2018_q4",
  "2018_q3",
  "2018_q2",
  "2018_q1",
  "2017_q4",
  "2017_q3",
  "2017_q2",
  "2017_q1",
  "2016_q4",
  "2016_q3",
  "2016_q2",
  "2016_q1",
  "2015_q4",
  "2015_q3",
  "2015_q2",
  "2015_q1",
  "2014_q4"
)
```

## Joining events with other tables

```{r}
nyc_events_joined <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_categories,
    by = "EvtId",
    all.x = TRUE,
    all.y = TRUE) %>%
  merge(
    y = nyc_event_organizers,
    by = "EvtId",
    all.x = TRUE,
    all.y = TRUE) %>%
  merge(
    y = nyc_event_listing_locations,
    by = "EvtId",
    all.x = TRUE,
    all.y = TRUE)

nyc_events_left_joined <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_categories, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_organizers, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_listing_locations, 
    by = "EvtId", 
    all.x = TRUE)
```


```{r cleaning for interactive event category monthly counts}
# Event category monthly counts over time, used for interactive component
nyc_events_category_counts_monthly <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_categories,
    by = "EvtId",
    all.x = TRUE) %>%
  filter(Dt < as.Date("2019-01-01")) %>%
  drop_na(Cat) %>%
  mutate(month = month(Dt, label=TRUE, abbr=TRUE), 
         year=year(Dt), 
         category = Cat, 
         date = as.yearmon(Dt)) %>%
  group_by(category, month, year, date) %>%
  summarize(count = n()) 

meaningful_timeseries_categories <- nyc_events_category_counts_monthly %>%
  group_by(category) %>%
  summarize(count = n()) %>%
  filter(count > 12) # Limit to categories with at least 12 distinct months of active event listings

nyc_events_category_counts_monthly %>%
  filter(category %in% meaningful_timeseries_categories$category) %>%
  write.csv("data/derived_data/monthly_category_counts.csv", row.names=FALSE)
```
```{r}
nyc_events_organizer_counts_monthly <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_organizers,
    by = "EvtId",
    all.x = TRUE) %>%
  filter(Dt < as.Date("2019-01-01")) %>%
  drop_na(Org) %>%
  mutate(month = month(Dt, label=TRUE, abbr=TRUE), 
         year=year(Dt), 
         organizer = Org, 
         date = as.yearmon(Dt)) %>%
  group_by(organizer, month, year, date) %>%
  summarize(count = n()) 

meaningful_timeseries_organizers <- nyc_events_organizer_counts_monthly %>%
  group_by(organizer) %>%
  summarize(count = n()) %>%
  top_n(30) 

nyc_events_organizer_counts_monthly %>%
  filter(organizer %in% meaningful_timeseries_organizers$organizer) %>%
  write.csv("data/derived_data/monthly_organizer_counts.csv", row.names=FALSE)
```

(TOOD: No longer accurate)
The events listing data is the main data table, therefore, we want to keep all rows in the events listing table even if there is no matching row in the other tables. As a result, we are performing left outer joins and keeping all rows of the `nyc_event_listing_events` data frame.


```{r events join categories}
nyc_events_with_categories <- 
  merge(x = nyc_event_listing_events, 
        y = nyc_event_categories, 
        by = "EvtId", 
        all.x = TRUE)

nyc_events_with_category_counts <- nyc_event_categories %>%
  drop_na() %>%
  group_by(EvtId) %>%
  summarize(CatCt = n()) %>%
  ungroup()
nyc_events_with_category_counts
```


```{r}
nyc_events_with_categories %>%
  group_by(Cat) %>%
  summarize(count=n())
```

```{r}
nyc_events_with_organizer_counts <- nyc_event_organizers %>%
  drop_na() %>%
  group_by(EvtId) %>%
  summarize(OrgCt = n()) %>%
  ungroup()
nyc_events_with_organizer_counts
```

```{r events with organizer and category counts}
nyc_events_with_cat_org_counts <- nyc_event_listing_events %>%
  merge(y = nyc_events_with_category_counts, 
        by = "EvtId", 
        all.x = TRUE, all.y = TRUE) %>%
  merge(y = nyc_events_with_organizer_counts, 
        by = "EvtId", 
        all.x = TRUE, all.y = TRUE)
```




```{r scratch work}
test <- nyc_event_organizers %>%
  mutate(Org = str_sub(Org, 1, 30)) %>%
  group_by(Org) %>%
  summarize(count=n()) %>%
  drop_na()
test
```

```{r}
library(stringdist)

test_grp <- test %>%
  mutate(grp = 
           cutree(hclust(stringdistmatrix(test$Org, method = "lv")), 
                  h = 15)) %>%
  group_by(grp) 

```

```{r}
# Does not work very well
test_grp %>%
  arrange(grp, count)
```
