# Missing values

```{r}
library(tidyverse)
source("projectlibs/missing_value_plot.R")
```

## Missing Values By Individual Tables

We first analyze the missing values pattern for all of our tables individually.

### Missing Values for Main Events Listing Table

As shown in the graph below, the majority of the data points have missing fields. The most missed fields are location_description (`LcDs`) and  cost_description (`CsDs`). 75% of the data points don't have those fields. We also have a lot of Email (`Em`) and Phone (`Ph`) missings (around 25% and 20% respectively). Those fields are not useful to our further analysis. For the description fields, they are textual and semantic, which would be interesting subjects for analysis and learning. However, the complexity of such analysis is beyond this study. As for contact information like email and phone number, they are not relevant to our analysis.

```{r nyc event listing events missing data}

missing_value_plot(nyc_event_listing_events, percent=TRUE)
missing_value_plot(nyc_event_listing_events, percent=FALSE)
col_missing = colSums(is.na(nyc_event_listing_events)) %>% sort(decreasing = TRUE)
col_missing
```
In the chart below, we dropped the fields mentioned above. As shown in the graph, almost all the data points are complete cases now. The few missing data in snippet (`Snp`), description (`Ds`) and url (`Url`) fields are not needed in our study. 

```{r nyc event listing events missing data pruned}
drop <- c("LcDs", "CsDs", "Ph", "Em")
nyc_envent_listing_events_sub <- nyc_event_listing_events[,!(names(nyc_event_listing_events) %in% drop)]

missing_value_plot(nyc_envent_listing_events_sub, percent=TRUE)
missing_value_plot(nyc_envent_listing_events_sub, percent=FALSE)
col_missing = colSums(is.na(nyc_envent_listing_events_sub)) %>% sort(decreasing = TRUE)
col_missing
```

### Missing Values for Auxiliary Tables

We also inspect the other auxiliary data tables that we are planning to use for our analysis.

```{r}
missing_value_plot(nyc_event_categories, percent = TRUE)
missing_value_plot(nyc_event_organizers, percent = TRUE)
colSums(is.na(nyc_event_organizers)) %>% sort(decreasing = TRUE)
```

As can be clearly seen from the missing data plots, the event category table itself does not have any missing data, and the event organizer table has very few missing organizer rows, more precisely, it only has 1 row of organizer missing.

The event id (`EvtId`) appears to be a unique identifier for events, and is always present in all rows of all tables we analyze.

As for the location data for the park events, as we can see from the char below, there are decent amount of missing fields in the data set. The most unfortunate missing field is zip (`Zip`), which can help us allocate the parks into smaller regions for further location-based studies. Since we have around 30% of the zip code data missing, the only location information for us to perform fine-grained regionalization of the parks is the longitude (`Long`) and latitude (`Lat`) field pair, in which we have 0 missing data. But the complexity would be much higher. Other missing data in address (`Addr`), accessibility (`Accs`) and park_id (`pkId`) are not study subjects in our project. It is interesting though, that the data set has complete longitude and latitude records for the events while misses 30% of the zip code data. With zip code being the apparent easier location data to acquire, we suspect that the longitude and latitude data are required for event registration so that it could be located on a map instantly. 
```{r nyc event listing locations missing data}
missing_value_plot(nyc_event_listing_locations, percent=TRUE)
missing_value_plot(nyc_event_listing_locations, percent=FALSE)
col_missing = colSums(is.na(nyc_event_listing_locations)) %>% sort(decreasing = TRUE)
col_missing
```
After dropping "Addr", "Accs", "pkId", "Zip" columns in the data set, we can have a closer look at the missing data percentage for Borough (`Bo`) and Location Name (`LcNm`). Location name is not a reliable source of information for location identification, since locations are sketchy and not rigidly defined. The missing data in the Borough (`Bo`) fields are minimal, whish would not affect our coarse-grained location-based studies. 
```{r nyc event listing locations missing data w/ dropped columns}
drop <- c("Addr", "Accs", "pkId", "Zip")
nyc_event_listing_locations_sub <- nyc_event_listing_locations[,!(names(nyc_event_listing_locations) %in% drop)]
missing_value_plot(nyc_event_listing_locations_sub, percent=TRUE)
missing_value_plot(nyc_event_listing_locations_sub, percent=FALSE)
col_missing = colSums(is.na(nyc_event_listing_locations_sub)) %>% sort(decreasing = TRUE)
col_missing
```

How about the missing patterns when joining the individual additional information tables (categories, locations, and organizers) with the main events listing table? We will explore in the following section.



## Missing Values for Joined Data Table

Next, we analyze the missing value patterns when we join our main events listing table with the auxiliary tables.

### Using Full Outer Joins

As previously discussed, the location description (`LcDs`), cost description (`CsDs`), email (`Em`), phone (`Ph`),
snippet (`Snp`), description (`Ds`), url (`url`) are not interesting variables for the purpose of our project, as 
most of them are free-form text data, thus we are dropping them in this section for analyzing the missing patterns
for the events listing data joined with other additional information tables since we do not really care if they are missing or not. 

Also, as disccused previously, title
(`Title`), is cost free (`CsFr`), date (`Dt`), end time (`ET`), is must see (`MtSe`), notice (`Ntc`), start time
(`ST`) all have no missing data in the main events listing data table, we will only keep title (`Title`) to
represent these variables in this section.

As analyzed above, `Lat` / `Long` are always present in the locations table, thus we will use `Lat` to count the 
number of unique locations per `EvtId`, thus we used it to determine if location information is missing or not in 
the joined table.


```{r}
nyc_events_with_is_na <- nyc_events_joined[c("EvtId", "Title", "Lat", "Org", "Cat")] %>%
  group_by(EvtId) %>%
  summarize(
    Title = sum(!is.na(Title)),
    Org = sum(!is.na(Org)),
    Cat = sum(!is.na(Cat)),
    Loc = sum(!is.na(Lat))
  ) %>%
  mutate_all(na_if, 0) %>%
  ungroup()

missing_value_plot(nyc_events_with_is_na, percent = TRUE)
colSums(is.na(nyc_events_with_is_na)) %>% sort(decreasing = TRUE)
```


When only looking at these variables, almost 75% of the rows have complete information. 

There are roughly 15% cases from the joined dataset that have all of the event_id, organizer, category, and
location information, but missing the title (and thus the other variables information) only, meaning that there
are close to 15% of the events with an event_id that appear in all of the auxiliary tables but not the main event 
listing table. Therefore, if we performed left outer joins on the tables (where we keep the main event listing 
table on the left), we should be getting close to 90% of complete cases.

Also, roughly more than 20% of the cases are missing title, these are the events with event_id appear in at least one of the auxiliary tables but not in the main table.

It is also interesting to see that there are no cases of all organizer, category, location information missing but with title information present. So any listed events in the main dataset have at least one of the three information (organizer, category, location).

A potential explanation for the above facts might be that a placeholder event can be registered in the 
system with a minimum one of these three pieces of information and an event_id was generated, but the event
may or may not ultimately get listed and stored in the main event listing table.


### Using Left Outer Joins

The following plot demonstrates the missing pattern analysis when we instead perform left joins of the main table 
with the auxiliary tables. It confirms that we have close to 90% (from the plots, more than 87.5%) of complete 
cases among these variables, and the top missing variable is organizer (less than 12.5%), followed by category, 
followed by location information. 


```{r}
nyc_events_left_joined <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_categories, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_organizers, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_locations, 
    by = "EvtId", 
    all.x = TRUE)

nyc_events_left_joined[c("EvtId", "Title", "Lat", "Org", "Cat")] %>%
  group_by(EvtId) %>%
  summarize(
    Title = sum(!is.na(Title)),
    Org = sum(!is.na(Org)),
    Cat = sum(!is.na(Cat)),
    Loc = sum(!is.na(Lat))
  ) %>%
  mutate_all(na_if, 0) %>%
  ungroup() %>%
  missing_value_plot(percent = TRUE)

```

As a result of the analysis in this section, we will use the table produced by the left outer join (including the 
variables that were left out in this section due to the `Title` representation) for our main 
analysis in later chapters, since it includes the most complete and interesting information.


## Top Categories with Missing Organizer

It is still pretty surprising to us that after the left outer join, we still have more than 10% of the cases 
missing an organizer, as we thought all events should have an organizer. Therefore, we wanted to see what are the top categories of events without the organizer information.

Note here that an event might be associated with multiple categories, so the total number of missing organizers 
across the different categories would be higher than the number of events with missing organizers in the previous 
section.


```{r}
nyc_events_left_joined %>%
  group_by(Cat) %>%
  summarize(Num_Missing_Org = sum(is.na(Org))) %>%
  ungroup() %>%
  arrange(desc(Num_Missing_Org)) %>%
  top_n(20) %>%
  ggplot(aes(x = Num_Missing_Org, y = fct_reorder(Cat, Num_Missing_Org))) + 
  geom_point(color = "blue") +
  ggtitle("Top event categories missing an organizer") +
  labs(x = "Number missing an organizer", y = "Event Category") +
  theme_linedraw() 
```

So it seems that sports-related events (with Fitness/Running/Sports/Outdoor Fitness categories) are the most common themes  for events missing organizer information in our dataset. A closer look at the specific event titles suggests that  a lot of them were running events organized by "NYRR" in different parks:

```{r}
nyc_events_left_joined %>%
  group_by(Title) %>%
  summarize(Num_Missing_Org = sum(is.na(Org))) %>%
  ungroup() %>%
  arrange(desc(Num_Missing_Org)) %>%
  top_n(20) %>%
  ggplot(aes(x = Num_Missing_Org, y = fct_reorder(Title, Num_Missing_Org))) + 
  geom_point(color = "blue") +
  ggtitle("Top event categories missing an organizer") +
  labs(x = "Number missing an organizer", y = "Event Title") +
  theme_linedraw() 
```

## Missing Values for Crimes in Parks

```{r nyc park crimes missing data}
parks <- unique(nyc_parks_crime$Park)

park_qtr_missing_check <- data.frame(
  rep(parks, each = length(crime_quarters)), 
  rep(crime_quarters, times = length(parks))
  )
names(park_qtr_missing_check) <- c("Park", "Qtr")
nyc_parks_crime_complete_quarters <- merge(nyc_parks_crime, park_qtr_missing_check, by=c("Park", "Qtr"), all.y = TRUE)

missing_value_plot(nyc_parks_crime_complete_quarters, percent=TRUE)
col_missing = colSums(is.na(nyc_parks_crime_complete_quarters)) %>% sort(decreasing = TRUE)
col_missing
```
The NYC park crime data originally comes in different files by quarter. We consolidated all of them to analyze the missing data pattern. In order to ensure that no parks are missed in any of the data files, we created a data set with all the <park, quarter> pairs there could possibly be in the crime data set. Then we merge the new data set with the crime data set to perform the missing data analysis. As shown in the graph above, there is indeed no missing data in the crime data set. 
