# Results

```{r}
library(ggmap)
library(tidyverse)
library(leaflet)
```

Top Event Categories

```{r}
top_categories <- nyc_event_categories %>%
  group_by(Cat) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(20)

top_categories
```

## Events and Parks Analysis

### Most lively parks
```{r}
nyc_events_joined %>%
  drop_na(EvtId, Lat, Long) %>%
  mutate(gps_location=paste(Lat,Long, sep=",")) %>%
  group_by(gps_location) %>%
  summarise(num_events=n()) %>%
  arrange(desc(num_events)) %>%
  top_n(10) %>%
  ggplot(aes(x = num_events, y = fct_reorder(gps_location, num_events))) +
  geom_point(color = "blue") +
  ggtitle("Most lively parks") +
  labs(x = "Number of events", y = "Park Location") +
  theme_linedraw()
```

### Most Diversified Park
```{r}
nyc_events_joined %>%
  drop_na(EvtId, Lat, Long, Cat) %>%
  mutate(gps_location=paste(Lat,Long, sep=",")) %>%
  group_by(gps_location) %>%
  summarise(num_cats=length(unique(Cat))) %>%
  arrange(desc(num_cats)) %>%
  top_n(10) %>%
  ggplot(aes(x = num_cats, y = fct_reorder(gps_location, num_cats))) +
  geom_point(color = "blue") +
  ggtitle("Most diversified parks") +
  labs(x = "Number of event categories", y = "Park Location") +
  theme_linedraw()
```

### Ratio of free events in parks
```{r}
nyc_events_joined %>%
  drop_na(EvtId, Lat, Long, CsFr) %>%
  mutate(gps_location=paste(Lat,Long, sep=",")) %>%
  group_by(gps_location) %>%
  #filter(n() > 100) %>%
  summarise(free_ratio=sum(CsFr)/n()) %>%
  ggplot(aes(free_ratio)) +
  geom_histogram(color = "blue", fill = "lightblue", binwidth=0.1) +
  ggtitle("Histogram of free event ratio")
```


## Spatial Analysis

### Cost

From the missing values chapter, we do not have missing data for `CsFr`, `Lat`, `Long`

```{r}

nyc_events_fee_location <- nyc_event_listing_events[c("EvtId", "CsFr")] %>%
  merge(
    y = nyc_event_locations, 
    by = "EvtId", 
    all.x = TRUE) %>% 
  .[c("EvtId", "CsFr", "Lat", "Long")] %>%
  group_by(CsFr, Lat, Long) %>%
  summarize(count = n()) %>%
  ungroup()

# qmplot(Long, Lat, data = test, maptype = "toner-lite", color = I("red"))
```


TODO Too much data for this, select a year maybe

```{r}
cost_palette <- colorFactor(palette='RdBu', domain=nyc_events_fee_location$CsFr)

leaflet(nyc_events_fee_location) %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 10) %>%
  addCircleMarkers(~Long, ~Lat, radius = 5, stroke=FALSE, fillOpacity = 0.5, fillColor = ~ cost_palette(CsFr)) %>%
  addLegend("topright", pal = cost_palette, values = nyc_events_fee_location$CsFr, title = "Is Free")
```



### By Category

```{r}

test_cats <- c("Waterfront", "Games", "Festivals")

top_category_events_locations <- nyc_event_listing_events[c("EvtId")] %>%
  merge(
    y = nyc_event_categories, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_locations, 
    by = "EvtId", 
    all.x = TRUE) %>% 
  .[c("EvtId", "Cat", "Lat", "Long")] %>%
  filter(Cat %in% test_cats)
  #filter(Cat %in% top_categories$Cat)

categories_palette <- colorFactor(palette='Set1', domain=top_category_events_locations$Cat)
  
leaflet(top_category_events_locations) %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 10) %>%
  addCircleMarkers(~Long, ~Lat, radius = 5, stroke=FALSE, fillOpacity = 0.5, fillColor = ~ categories_palette(Cat)) %>%
  addLegend("topright", pal = categories_palette, values = top_category_events_locations$Cat, title = "Category")
```


### By Organizer


```{r}
top_organizers <- nyc_event_organizers %>%
  group_by(Org) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(8)

top_organizers

top_orgs_events_locations <- nyc_event_listing_events[c("EvtId")] %>%
  merge(
    y = nyc_event_organizers, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_locations, 
    by = "EvtId", 
    all.x = TRUE) %>% 
  .[c("EvtId", "Org", "Lat", "Long")] %>%
  filter(Org %in% top_organizers$Org)

organizers_palette <- colorFactor(palette='Set1', domain=top_orgs_events_locations$Org)
  
leaflet(top_orgs_events_locations) %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 10) %>%
  addCircleMarkers(~Long, ~Lat, radius = 5, stroke=FALSE, fillOpacity = 0.5, fillColor = ~ organizers_palette(Org)) %>%
  addLegend("topright", pal = organizers_palette, values = top_orgs_events_locations$Org, title = "Organizer")
```


## References

D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2. The R Journal, 5(1), 144-161. URL
  http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf

