# Results

```{r}
library(parcoords)
library(GGally)
library(leaflet)
library(lubridate)
library(vcd)
library(vcdExtra)
library(RColorBrewer)
library(tidyverse)
library(ggrepel)
```

## Events and Parks Analysis


```{r Events and Parks Analysis data prep}
top_categories <- nyc_event_categories %>%
  group_by(Cat) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(30)

top_categories_levels <- levels(fct_reorder(top_categories$Cat, top_categories$count))

top_locations <- nyc_event_listing_locations %>%
  group_by(LcNm) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(30)
```


### Most Diverse Park

Since our main dataset is about the events in the different parks in NYC, very naturally, the first thing we'll look at is what are some of the parks and locations with the most events:

```{r Most Diverse Park Cleveland Dot Plot}
ggplot(top_locations, aes(x = count, y = fct_reorder(LcNm, count))) +
  geom_point(color = "blue") +
  ggtitle("Top Event Locations") +
  labs(x = "Number of events", y = "Event Location") +
  theme_linedraw()
```

From this Cleveland Dot Plot of top 30 parks with the most number of events listed from 2013-2019 in our dataset, 
we can see that the top 2 locations are prety much outliers from the rest of the group, with "Poe Park Visitor 
Center" having more than 7000 events, followed by "Queens Botanical Garden", with more than 3000 events, and the 
rest of the top 30 locations all have fewer than 2000 events over the years.

After seeing that the top event locations are hosting several thousands of events over the years 2013-2019, we 
wonder which of them are the most diverse ones, in the sense that they have a lot of distinct event organizers 
organizing events of different categories in them.

```{r Most Diverse Park location scatter}

nyc_events_left_joined %>%
  drop_na(LcNm, Org, Cat) %>%
  group_by(LcNm) %>%
  summarize(UniqueCats = n_distinct(Cat), UniqueOrgs = n_distinct(Org)) %>%
  ggplot(aes(UniqueOrgs, UniqueCats)) +
  geom_point(alpha = 0.4, size = 2, stroke = 0) +
  geom_text(aes(label = ifelse(
  (UniqueOrgs > 24) | 
  (UniqueCats > 45) | 
  (LcNm == "Poe Park Visitor Center") |
  (LcNm == "Queens Botanical Garden"), as.character(LcNm),'')), 
            size = 2.3,
            hjust = 0.82, 
            vjust = 0.9) +
  labs(x = "Count of Distinct Organizers", y = "Count of Distinct Event Categories") + 
  ggtitle("Parks Hosting Different Event Categories Organized by Event Organizers") + 
  theme(panel.background = element_blank())
```

We have labeled the locations with more than 24 unique organizers who have hosted events, or have events of more 
than 45 different event categories over the years, as well as the locations with most events - ""Poe Park Visitor 
Center" and "Queens Botanical Garden" - interesting enough that for these two, we can notice although they have a 
diverse number of event categories, they both had no more than 10 distinct organizers who have organized events 
in them.

Surprisingly, the most famous parks like "Central Park" and "Washington Square Park" do not show up. From the 
location names, we realize these parks might be too large and have different sub locations (e.g. "Visitor Center" 
inside of "Conference House Park") listed as the location name in the datasets we have. We then again plot the 
scatterplot using `PkId`:

```{r Most Diverse Park park id scatter}
parks_pkid_diversity <- nyc_events_left_joined %>%
  drop_na(PkId, Org, Cat) %>%
  group_by(PkId) %>%
  summarize(UniqueCats = n_distinct(Cat), UniqueOrgs = n_distinct(Org))

ggplot(parks_pkid_diversity, aes(UniqueOrgs, UniqueCats)) +
  geom_point(alpha = 0.4, size = 2, stroke = 0) +
  geom_text(aes(label = ifelse((UniqueOrgs > 40) | (UniqueCats > 60), as.character(PkId),'')), 
            size = 3,
            hjust = 1, 
            vjust = 1) +
  labs(x = "Count of Distinct Organizers", y = "Count of Distinct Event Categories") + 
  ggtitle("Parks Hosting Different Event Categories Organized by Event Organizers") + 
  theme(panel.background = element_blank())

```

A quick Googling using the longitude and latitude of these park ids tells us that, the labels from left to right are:

```{r Most Diverse Park mapping table}
knitr::kable(
  tibble(
    "Park Id" = c(
      "X039",
      "B073",
      "X104",
      "X092",
      "X045",
      "M029",
      "R006",
      "M098",
      "M042",
      "Q099",
      "M010"
    ),
    "Park Name" = c(
      "Pelham Bay Park",
      "Prospect Park",
      "Williamsbridge Oval Skate Park",
      "Van Cortlandt Park",
      "St. Mary's Park",
      "Fort Tryon Park",
      "Conference House Park",
      "Washington Square Park",
      "Inwood Hill Park",
      "Flushing Meadows Corona Park",
      "Central Park"
    )
  ),
  booktabs = TRUE,
  caption = 'Park Id to Park Name Mapping for Diverse Parks'
)
```


all of which are more well-known parks in NYC.

From this scatter plot, it is certainly apparent that there seems to be a correlation that the parks with more 
diverse groups of organizers choosing them as event locations tend to have events with more diverse categories as 
well, which makes us wonder if different organizers have preferences over certain event categories, for which we 
will analyze in the next section.

### Organizer Preferences on Categories and Locations

```{r Organizer Preferences on Categories and Locations scatter}

nyc_events_left_joined %>%
  drop_na(PkId, Org, Cat) %>%
  group_by(Org) %>%
  summarize(UniqueCats = n_distinct(Cat), UniquePkIds = n_distinct(PkId)) %>%
  ggplot(aes(UniquePkIds, UniqueCats) ) +
  geom_text_repel(aes(label = ifelse((UniquePkIds > 50) | (UniqueCats > 50), as.character(Org),'')),
            size = 3,
            max.overlaps = 30) +
  geom_point(alpha = 0.4, size = 2, stroke = 0) +
  labs(x = "Count of Distinct Parks", y = "Count of Distinct Event Categories") + 
  ggtitle("Organizers Organizing Different Event Categories Hosted in Different Parks") + 
  theme(panel.background = element_blank())

```


### Cost, Borough, Duration, and Time of Events

```{r Cost, Borough, Duration, and Time of Events box}
nyc_events_with_duration <- nyc_events_left_joined %>%
  drop_na(Dt, ET, ST, Cat) %>%
  mutate(Duration=as.numeric(difftime(
    strptime(paste(Dt, ET), "%Y-%m-%d %H:%M"), strptime(paste(Dt, ST), "%Y-%m-%d %H:%M"), units = "hours"))) %>%
  filter(Duration > 0) %>%
  select(c('Cat', 'Duration'))
top_category <- nyc_events_left_joined %>%
  group_by(Cat) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  top_n(10) %>% select(c('Cat'))
nyc_events_with_duration %>%
  merge(top_category, by = c('Cat'), all.y=TRUE) %>%
  ggplot2::ggplot(aes(Duration, reorder(Cat,-Duration, median))) + 
    geom_boxplot() + 
    labs(title="Boxplots of event duration\nfor top 10 event categories") +
    ylab("Category") + 
    theme_grey(14)
```


```{r Cost, Borough, Duration, and Time of Events pairs, fig.width=9, fig.height=9}
get_duration_category <- function(Dt, ST, ET) {
  d <- as.numeric(difftime(strptime(paste(Dt, ET), "%Y-%m-%d %H:%M"), strptime(paste(Dt, ST), "%Y-%m-%d %H:%M")))/3600
  case_when (
    d < 1 ~ "< 1",
    (d >= 1.00 & d < 3.00) ~ "1-3",
    (d >= 3.00 & d < 5.00) ~ "3-5",
    d >= 5.00 ~ ">= 5"
  )
}
get_time_of_day <- function(ST){
  hour = ST %>% map(~ as.numeric(strsplit(.x, ':')[[1]][1]))
  case_when(
   hour >= 0 & hour < 6 ~ "Night",
   hour >= 6 & hour < 12 ~ "Morning",
   hour >= 12 & hour < 18 ~ "Afternoon",
   hour >= 18 & hour <24 ~ "Evening"
 )
}

free_mosaic <- nyc_events_left_joined %>%
  drop_na(Bo, CsFr, Dt, ST, ET) %>%
  mutate(Duration=get_duration_category(Dt, ST, ET),
         TimeofDay=get_time_of_day(ST))

free_mosaic$CostFree = as.factor(ifelse(free_mosaic$CsFr == 0, "Not Free", "Free"))
free_mosaic$Borough = as.factor(free_mosaic$Bo)
free_mosaic$Duration = fct_relevel(free_mosaic$Duration, levels = c("< 1", "1-3", "3-5", ">=5"))
free_mosaic$Time = fct_relevel(free_mosaic$TimeofDay, levels = c("Morning", "Afternoon", "Evening", "Night"))

mosaic_select <- free_mosaic %>%
  select(CostFree, Borough, Duration, Time)

pairs(table(mosaic_select),
      highlighting=2, 
      diag_panel = pairs_diagonal_text(
        distribute = "margin",
        rot = 45)
      )
```



## Timeseries Analysis

Now, it is natural to think that there is some seasonality in the events that happen in the NYC public locations.
Here we are going to perform an analysis on what are some interesting seasonal patterns

### Timeseries Analysis on NYC Parks Events

First, we simply take an overview look at the count of listed events over time:

```{r Timeseries Analysis on NYC Parks Events line}
nyc_events_timeseries <- nyc_event_listing_events %>%
  group_by(Dt) %>%
  summarize(Count = n()) %>%
  ungroup()

ggplot(nyc_events_timeseries, aes(Dt, Count)) + 
  geom_line() + 
  labs(x = "Date", y = "Count of Events") +
  ggtitle("Count of NYC Events in Parks Over Time")
```

Here, we observe a few things at first sight:

* The data counts appears to be abnormally low after the start of 2019, probably due to limited data availablility in the dataset, therefore, we are going to only use data prior to 2019 in our following timeseries analysis on the events daata in this section.
* The seasonality pattern looks similar for each year.
* There is definitely fewer events in winters around Janurary and generally more events in the summers.

The previous plot, although helps us identify the general pattern, looks pretty noisy, so next let's take a different view, which is less noisy:

```{r Timeseries Analysis on NYC Parks Events less noisy, fig.width=10}
nyc_events_timeseries_weekly <- nyc_event_listing_events %>%
  filter(Dt < as.Date("2019-01-01")) %>%
  mutate(Month = month(Dt, label=TRUE, abbr=TRUE), Wday = wday(Dt, label = TRUE, abbr = TRUE), Year=year(Dt)) %>%
  group_by(Month, Wday, Year) %>%
  summarize(Count = n())

ggplot(nyc_events_timeseries_weekly, aes(Wday, Count)) +
  geom_bar(stat="identity", color = "blue", fill = "lightblue") +
  facet_grid(cols=vars(Month), rows=vars(Year))+
  ylim(0,500) + 
  ggtitle("Weekly Count of Park Events by Month") + 
  labs(x = "Weekday", y = "Count of Events") +
  ggtitle("Count of NYC Events in Parks By Weekday in Each Month") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
```

Here, we are aggregating the event counts by month, and plotted bar charts faceted by year. It is very clear from 
these charts that the monthly counts and trends do indeed have a very similar pattern across the years 2013-2018. 
Other than some fluctuations, January through March are typically the more quiet months with fewer than 500 events
in most years, then the number of events gradually increases in April, May and June, and peaks in July or August 
with more than 1500 events, and then gradually decreases into the winter season, with around 500 events in 
December.

Now, how about different days of week? It's certainly natural to speculate that there will be more events over 
weekends than during weekdays. Here, since we have concluded in the previous section that the seasonal trend is 
similar across the years, we take a closer look at the number of listed events by day of week in 2017 
to get an idea.

As expected, Saturdays do typically have a lot more events than the other days, especially during the summer 
season. It is somewhat surprising that Sundays only have about similar number of events as Tuesdays, Wednesdays 
and Thursdays. We speculate that for weekends, people prefer to stay at home with family and prepare for the new 
week on Sundays and prefer to go out to events on Saturdays, and therefore the event organizers have a preference 
to organize events on Saturdays. It is not surprising that Monday have the fewest events since people are 
typically busy on the first business day of the week. Fridays also appear to have the 2nd fewest number of events 
among the different days of week, as people wrap up the work for the week, and usually have corporate events or 
team bonding activities on Fridays.

### Timeseries Analysis on Event Categories

After the initial timeseries analysis, we feel that it's likely that other than the general seasonality and weekly
trends, different event sub-categories could have very different trends. Thus we will take a closer look at the top 30 categories for NYC parks events in this section.

```{r Timeseries Analysis on Event Categories data prep}
nyc_events_categories_timeseries_monthly <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_categories, 
    by = "EvtId", 
    all.x = TRUE) %>%
  filter(Dt < as.Date("2019-01-01")) %>%
  filter(Cat %in% top_categories$Cat) %>%
  mutate(Year = year(Dt), 
         Month = month(Dt, label = TRUE, abbr = TRUE), 
         DoW = wday(Dt, label = TRUE),
         Cat = fct_rev(factor(Cat, levels = top_categories_levels)))
```

First, we look at if there are trends in the numbers of events for different categories over the years.


```{r Timeseries Analysis on Event Categories yearly, fig.height=9}

nyc_events_categories_timeseries_monthly %>%
  group_by(Year, Cat) %>%
  summarize(Count = n()) %>%
  ungroup() %>%
  ggplot(aes(Year, Count)) +
  geom_col(color = "blue", fill = "lightblue") +
  facet_wrap(~Cat, ncol = 4) +
  ggtitle("Yearly Count of Events for top 30 Event Categories in NYC Parks") + 
  labs(x = "Year", y = "Count of Events") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```


We can notice on this plot that for most of the top categories, the number of events listed in each year from 2013
to 2018 more or less just fluctuates and stays mostly flat. However, for category "Education", there is a very 
noticeable increase for year 2018, and for categories "Seniors", "Accessible", "Film", "Dance", "Free Summer 
Movies", there appear to be an uptrend, whereas for categories like "Games" and "Birding", there appear to be a 
downtrend.

Next, we look at if all categories follow the general seasonality discussed in the previous section, or there are 
noticeable differences in the different categories.


```{r Timeseries Analysis on Event Categories monthly, fig.height=9}
nyc_events_categories_timeseries_monthly %>%
  group_by(Month, Cat) %>%
  summarize(Count = n()) %>%
  ungroup() %>%
  ggplot(aes(Month, Count)) +
  geom_col(color = "blue", fill = "lightblue") +
  scale_x_discrete(breaks=c("Mar", "Jun", "Sep", "Dec"),) + 
  facet_wrap(~Cat, ncol = 4) +
  ggtitle("Monthly Count of Events for top 30 Event Categories in NYC Parks") + 
  labs(x = "Month", y = "Count of Events") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```


Here, we can see that yes, for most of the top 30 categories, they mostly follow the "more events in the summer, 
fewer in the winter" general rule, with some other interesting observations, most noticeably:

* "Nature" events appear to be relatively flat from April through October, which is Spring through Autumn for NYC, then reduces during winter.
* "Arts & Crafts", "Poe Park Visitor Center", "Urban Park Rangers", and "Talks" events look to be relatively flat and just fluctuates throughout the year, with no apparent peaks.
* "Volunteer", "It's My Park" and "City Parks Foundation" events appear to have peaks in May and October.
* "History", "Tours", and "Historic House Trust Sites" events appear to increase from Januray through October, peaks at October, and then decreases in November and December.
* "Birding" has peaks in April and May.

Lastly, we also inspect the day of week pattern of when the top 30 event categories happen.

```{r Timeseries Analysis on Event Categories day of week, fig.height=9}
nyc_events_categories_timeseries_monthly %>%
  group_by(DoW, Cat) %>%
  summarize(Count = n()) %>%
  ungroup() %>%
  ggplot(aes(DoW, Count), color = "blue", fill = "lightblue") +
  geom_col(color = "blue", fill = "lightblue") +
  facet_wrap(~Cat, ncol = 4) +
  ggtitle("Day of Week Count of Events for top 30 Event Categories in NYC Parks") + 
  labs(x = "Day of Week", y = "Count of Events") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

As we can see from this plot, most of the top 30 event categories again follow the general "most events on Saturdays, fewest on Mondays" rule, again with some other interesting observations, most noticeably:

* "Seniors" has fewest events on Sundays and most events on Tuesdays, and relatively about the same count of events for different days of the week.
* "Volunteer", "It's My Park", "Urban Park Rangers", and "Birding" events happen primarily over weekends, and have very few events during weekdays.
* "Accessible", "Sports", and "Games" events happen about the same number of times during different days of the week.
* "Film" and "Free Summer Movies" have peaks on Fridays.
* "City Parks Foundation" events happen more frequently during weekdays than during weekends.
* "Poe Park Visitor Center" events do not occur on Sundays and Mondays.

## Crimes in Parks

Now we take a look at the crimes in the NYC parks recorded by NYPD.

### Crimes in Parks Overview

```{r Crimes in Parks Overview}
fills7 <- brewer.pal(7, 'Accent')

nyc_parks_crimes_total <- nyc_parks_crime %>%
  group_by(Boro) %>%
  summarize(Mdr = sum(Mdr), 
            Rape = sum(Rape), 
            Rob = sum(Rob), 
            Aslt = sum(Ass), 
            Burg = sum(Burg), 
            GL = sum(G_L), 
            GLM = sum(GL_M_V)) %>%
  pivot_longer(!Boro, names_to = "CrimeType", values_to = "Freq") %>%
  mutate(CrimeType = fct_relevel(CrimeType, "GLM", "Rob", "Rape", "Mdr", after = Inf)) %>%
  mutate(Boro = fct_recode(Boro, 
                           B = "BROOKLYN", 
                           M = "MANHATTAN", 
                           R = "STATEN ISLAND", 
                           Q = "QUEENS", 
                           X = "BRONX", 
                           "B/Q" = "BROOKLYN/QUEENS")) %>%
  mutate(Boro = fct_relevel(Boro, "B/Q"))

vcd::mosaic(CrimeType ~ Boro, 
            nyc_parks_crimes_total, 
            direction = c("v", "h"),
            labeling = labeling_border(tl_labels = c(FALSE, FALSE),
                                       gp_labels = gpar(fontsize = 9),
                                       set_varnames = c(Boro = "Borough", CrimeType = "Crime Type"),
                                       rot_labels = c(0, 0, 0, 0)),
            highlighting_fill = fills7,
            main = "Mosaic Plot of Crime Types in Boroughs of NYC")
```

The distribution of crime types in the different Boroughs of NYC over the years of 2014 through 2021 is actually different, especially around "Grand Larceny" and "Robbery" - for example, Manhattan has more "Grand Larceny" than "Robbery", whereas Staten Island and Bronx both have more "Robbery" than "Grand Larceny".

### Crime Timeseries Analysis
```{r Crime Timeseries Analysis, fig.width=9}
nyc_crimes_timeseries_quarterly <- nyc_parks_crime %>%
  drop_na(Boro, Qtr) %>%
  group_by(Boro, Qtr) %>%
  summarize(Mdr = sum(Mdr), 
            Rape = sum(Rape), 
            Rob = sum(Rob), 
            Aslt = sum(Ass), 
            Burg = sum(Burg), 
            GL = sum(G_L), 
            GLM = sum(GL_M_V),
            Tot = sum(Tot)) %>%
  pivot_longer(cols=3:10, names_to="Crime", values_to="Count")
ggplot(nyc_crimes_timeseries_quarterly, aes(Qtr, Count, group=Boro, color=Boro)) +
  geom_line(size = 0.5) +
  facet_grid(rows=vars(Crime), scales="free_y")+
  ggtitle("Count of Park Crimes by Category") + 
  labs(x = "Quarter", y = "Count of Reports") +
  ggtitle("Count of NYC Park Crimes in Parks By Quarter") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
        # axis.text.y = element_text(size = 12),
        # axis.title = element_text(size = 14),
        # strip.text.y = element_text(size = 14),
       
```


### Crimes in parks and events in parks

It is challenging to join the two datasets entirely based on the parks, as they come from two different data sources and do not share identifier keys for the parks. However, it is possible to manually select a few parks from both datasets and inspect the park names and try to match and join the data for those selected parks.

```{r Crimes in parks and events in parks top crimes table}
top10_parks_most_crime <- nyc_parks_crime %>% 
  group_by(Park) %>%
  summarize(count_of_crimes = sum(Tot)) %>%
  arrange(desc(count_of_crimes)) %>%
  top_n(10)

top10_parks_most_events <- nyc_event_listing_events[c("EvtId")] %>%
  merge(
    y = nyc_event_listing_locations, 
    by = "EvtId", 
    all.x = TRUE) %>%
  drop_na(PkId) %>%
  group_by(PkId) %>%
  summarize(count_of_events = n()) %>%
  arrange(desc(count_of_events)) %>%
  top_n(10)

knitr::kable(
  list(top10_parks_most_crime, top10_parks_most_events),
  booktabs = TRUE,
  caption = 'Parks with most crimes and Parks with most events'
)
```

A quick data aggregation gives us the top 10 parks with most crimes and top 10 parks with the most events, as shown in the data table above. However, we discovered that somehow the crimes data do not have any records for crimes in "Central Park" (M010). Together with our interests of the most diverse parks as analyzed in the [Most Diverse Park] section, we will select the following parks to analyze, and use this manually created mapping as the helper table to join our data:

```{r Crimes in parks and events in parks mapping table}

park_crime_event_mapping <- tibble(
  PkId = c(
    "Q099",
    "B169",
    "M104",
    "M098",
    "B073",
    "M072",
    "M089",
    "X010",
    "M008",
    "X002",
    "X040",
    "X209",
    "R006",
    "B431",
    "M029",
    "X092",
    "X039",
    "X104",
    "X045",
    "M042"
  ),
  Park = c(
    "FLUSHING MEADOWS CORONA PARK",
    "CONEY ISLAND BEACH & BOARDWALK",
    "RANDALL'S ISLAND PARK",
    "WASHINGTON SQUARE PARK",
    "PROSPECT PARK",
    "RIVERSIDE PARK",
    "UNION SQUARE PARK",
    "CROTONA PARK",
    "BRYANT PARK",
    "BRONX PARK",
    "POE PARK",
    "WAVE HILL",
    "CONFERENCE HOUSE PARK",
    "BROOKLYN BRIDGE PARK",
    "FORT TRYON PARK",
    "VAN CORTLANDT PARK",
    "PELHAM BAY PARK",
    "WILLIAMSBRIDGE OVAL",
    "ST. MARY'S PARK BRONX",
    "INWOOD HILL PARK"
  ),
  "ParkName" = c(
    "Flushing Meadows Corona Park",
    "Coney Island Beach",
    "Randall's Island",
    "Washington Square Park",
    "Prospect Park",
    "Riverside Park",
    "Union Square Park",
    "Crotona Park",
    "Bryant Park",
    "Bronx Park",
    "Poe Park",
    "Wave Hill",
    "Conference House Park",
    "Brooklyn Bridge Park",
    "Fort Tryon Park",
    "Van Cortlandt Park",
    "Pelham Bay Park",
    "Williamsbridge Oval Skate Park",
    "St. Mary’s Park",
    "Inwood Hill Park"
  )
)

knitr::kable(
  park_crime_event_mapping,
  booktabs = TRUE,
  caption = 'Park in crime dataset to PkId in events dataset mapping to investigate'
)
```


We understand the limitations of picking only the parks with most crimes or most events or most diverse, however, 
it is very difficult for us to include parks with fewer events or crimes and link them from the two data sources, 
for many reasons, some of which were outlined in previous [Missing values] chapter (that 20% of our event listing 
data are missing `PkId`), and the [Most Diverse Park] section (that the same `PkId` includes many sub-areas 
of parks, so that for smaller or less popular parks, it is super difficult for us to map the names across the two 
datasets we have). Nevertheless, we still wanted to derive certain features of these parks and see what are some 
interesting patterns we can find. 

```{r Crimes in parks and events in parks pcp data prep}

interesting_park_crimes_mapped <- park_crime_event_mapping %>%
  merge(
    y = nyc_parks_crime, 
    by = "Park", 
    all.x = TRUE) %>%
  group_by(ParkName, Boro, Size) %>%
  summarize(ViolentCrm = sum(Mdr) + sum(Rape) + sum(Rob) + sum(Ass), 
            PropertyCrm = sum(Burg) + sum(G_L) + sum(GL_M_V),
            TotCrm = sum(Tot))

interesting_parks_events_mapped_summary <- nyc_event_listing_events %>%
  merge(
    y = nyc_event_listing_locations, 
    by = "EvtId", 
    all.x = TRUE) %>%
  drop_na(PkId) %>%
  merge(
    y = park_crime_event_mapping, 
    by = "PkId", 
    all.y = TRUE) %>%
  mutate(Duration=difftime(
    strptime(paste(Dt, ET), "%Y-%m-%d %H:%M"), strptime(paste(Dt, ST), "%Y-%m-%d %H:%M"), units = "hours")) %>%
  filter(Duration > 0) %>%
  .[c("PkId", "EvtId", "Duration", "CsFr", "Park", "ParkName")] %>%
  group_by(ParkName) %>%
  summarize(EvtCnt = n(), AvgDur = sum(Duration)/n(), FrRatio = sum(CsFr)/n())


parks_events_crime_relation_tidy <- parks_pkid_diversity %>%
  merge(
    y = park_crime_event_mapping, 
    by = "PkId", 
    all.y = TRUE) %>%
  .[c("UniqueCats", "UniqueOrgs", "ParkName")] %>%
  merge(
    y = interesting_parks_events_mapped_summary,
    by = "ParkName"
  ) %>%
  merge(
    y = interesting_park_crimes_mapped,
    by = "ParkName"
  )


parks_events_crime_relation_tidy$AvgDur <- as.numeric(parks_events_crime_relation_tidy$AvgDur, units = "hours")
parks_events_crime_relation_tidy <- parks_events_crime_relation_tidy[, c(2:6, 8:11)]
parks_events_crime_relation_tidy <- parks_events_crime_relation_tidy[, c(6, 1:3, 9, 4, 7:8, 4, 5)]
```

```{r Crimes in parks and events in parks pcp}
ggparcoord(parks_events_crime_relation_tidy, 
           columns = 1:10, 
           scale = "std", 
           alphaLines = 1 / 3, 
           splineFactor = 10) +
  geom_vline(xintercept = 1:10, color = "lightblue") +
  ggtitle("Parallel Coordinates Plot for Events and Crimes in NYC Parks") +
  labs(x = "Variable", y = "Univariately Std Re-Scaled Ratio")
```

For this parallel coordinates plot, we have arranged the x-axis columns to highlight interesting patterns, and the "Average Duration" (`AvgDur`) has been drawn twice for this reason too. From the several interesting dimensions we have derived from our events and crimes data for the 20 parks mentioned, there are a few interesting observations we can notice:

1. Other than a few outliers parks, for the 20 parks we analyzed, a lot of them have similar sizes and have similar ratio of free events out of all events.
2. We can again observe the "diversity" correlation (similarly observed in [Most Diverse Park] section) in terms of categories and organizers
3. Out of the 20 parks analyzed, larger parks have better diversity in terms of event categories and organizers
4. It is somewhat surprising to see that the more diverse parks actually listed fewer events
5. There appear to be a negative association between the number of events in parks and the number of crimes in parks, as can be seen by the lines flowing in opposite directions and form "crosses" between the `EvtCnt` and `TotCrm` columns on the plot.
6. We can similarly observe a negative association between the average duration of events (`AvgDur`) and the number of crimes (`TotCrm`), (and for both violent crime types (`ViolentCrm`) and property crime types (`PropertyCrm`)) in the 20 parks. This fact, together with observation number 5, makes us suspect that crimes more frequently happen when there are not a lot of people/events in the parks.
7. Other than a few outliers, it appears that parks with more violent crime types tend to have more property crime types as well.

## Spatial Analysis

From the [Missing values] chapter, we do not have missing data for `Lat`, `Long`, which certainly prompted us to perform a detailed Spatial Analysis on the events in the NYC parks.


### Spatial Analysis of Events By Organizer

From our analysis in the [Organizer Preferences on Categories and Locations] section, there certainly appears to be preferences of different organizers on the parks they usually organize events in. As a result, we want to take a look at the top 8 event organizers and see where they are hosting most of their events:

```{r Spatial Analysis of Events By Organizer}
top_organizers <- nyc_event_organizers %>%
  group_by(Org) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(8)

top_orgs_events_locations <- nyc_event_listing_events[c("EvtId")] %>%
  merge(
    y = nyc_event_organizers, 
    by = "EvtId", 
    all.x = TRUE) %>%
  merge(
    y = nyc_event_listing_locations, 
    by = "EvtId", 
    all.x = TRUE) %>% 
  .[c("EvtId", "Org", "Lat", "Long")] %>%
  filter(Org %in% top_organizers$Org)

organizers_palette <- colorFactor(palette='Set1', domain=top_orgs_events_locations$Org)
  
leaflet(top_orgs_events_locations) %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 10) %>%
  addCircleMarkers(~Long, ~Lat, radius = 3, stroke=FALSE, fillOpacity = 0.5, fillColor = ~ organizers_palette(Org)) %>%
  addLegend("topright", pal = organizers_palette, values = top_orgs_events_locations$Org, title = "Organizer")
```



Not surprisingly, we can see the organizer preferences on location of events very clearly:

* "Bryant Park Corp." solely organizes events in midtown Manhattan near Bryant Park.
* "Central Park Conservancy" organizes all of its events in Central Park.
* "Poe Visitor Center" only has events in Poe Park.
* "Prospect Park Alliance" events are all in Prospect Park.
* "Summer on the Hudson" events happen in different parks but all along the Hudson river.
* Most of "Queens Botanical Garden" events are in Queens Botanical Garden but they also organize some events outside of the Queens Botanical Garden, but mostly in Queens.
* "NYC Parks" and "Urban Park Rangers" are the more diverse organizers out of the top 8 active organizers that they organize events in more locations throughout the whole NYC areas in all boroughs, as similarly identified in the [Organizer Preferences on Categories and Locations] section. We can also observe an interesting preference of "Urban Park Rangers" to have more events along waterfronts when compared to "NYC Parks".

It is certainly very tempting for us to analyze the spatial distribution of events for different event categories,
however, based on our previous analysis, we feel we have too many events categories to have a clean spatial plot 
to analyze, also, as analyzed in the [Most Diverse Park] section, many parks are having more than 20 distinct 
categories of events, it would certainly give us lots of over-plotting, thus we refrain ourselves from doing a 
deep dive of the spatial distribution of different event categories here. However, there are also interesting 
findings in our previous analysis in the [Cost, Borough, Duration, and Time of Events] section, we 
particularly feel it is very interesting to see the distribution of the events on the map based on their start time of the day.

### Spatial Analysis on Events Start Time of the Day

Here, we have included the following spatial graph of events colored by their start time of the day, note that 
due to the large amount of data, we have only included events in 2018 in this plot, but it is very representative 
of all the years:


```{r Spatial Analysis on Events Start Time of the Day}

spatial_events_w_tod_2018 <- nyc_event_listing_events[c("EvtId", "Dt", "ST")] %>%
  merge(
    y = nyc_event_listing_locations, 
    by = "EvtId", 
    all.x = TRUE) %>% 
  .[c("EvtId", "ST", "Lat", "Long", "Dt")] %>%
  drop_na(Dt, ST) %>%
  mutate(TimeOfDay = get_time_of_day(ST)) %>%
  drop_na(TimeOfDay) %>%
  filter(year(.$Dt) == 2018)


spatial_events_w_tod_2018$TimeOfDay <- fct_relevel(
  factor(spatial_events_w_tod_2018$TimeOfDay), 
  levels = c("Night", "Evening", "Afternoon", "Morning"))
```

```{r}
time_palette <- colorFactor(palette='viridis', domain=spatial_events_w_tod_2018$TimeOfDay)

leaflet(spatial_events_w_tod_2018) %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 10) %>%
  addCircleMarkers(
    ~Long, 
    ~Lat, 
    radius = 3, 
    stroke=FALSE, 
    fillOpacity = 0.5, 
    fillColor = ~ time_palette(TimeOfDay)) %>%
  addLegend("topright", 
            pal = time_palette, 
            values = spatial_events_w_tod_2018$TimeOfDay,
            title = "Event Start Time")
```



From the spatial plot here, it is easy to see that, as also analyzed in section 
[Cost, Borough, Duration, and Time of Events], Brooklyn certainly has a lower ratio of morning events and higher 
ratio of evening events when compared to the other boroughs. Manhattan has a pretty balanced diversity of 
morning/afternoon/evening events. Manhattan also certainly has more densely organized events.

It is also interesting but not surprising to see that there are more evening events along different waterfront 
parks - who doesn't enjoy the night waterfront city view? For the "inner" located parks, there are certainly more 
morning/afternoon events than evening events, which certainly explains why Bronx has the highest ratio of 
morning events and lowest ratio of evening events out of all boroughs. 


